Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges

### 2017-08-17 paper.tex
reviewed section 3.1
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

Bounded Model Checking (BMC) based on Boolean Satisfiability (SAT) was originally proposed to verify hardware designs~\cite{Biere99,handbook09}. Indeed, Biere {\it et al.} were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by standard SAT solvers. BMC based on SMT~\cite{BarrettSST09}, in turn, was originally proposed to deal with increasing software verification complexity~\cite{Armando2009}. 
%
In general, BMC techniques aim to check the violation of a given (safety) property at a given system depth, as shown in Figure~\ref{bounded-model-checking}. 
%
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{figure2.jpg}
	\caption{Bounded Model Checking.}
	\label{bounded-model-checking}
\end{figure}

@@

\noindent Here, $\phi_{k}$ represents a safety property $\phi$ in step $k$, $I$ is the set of initial states of $M$, $R\left(s_{i},s_{i+1}\right)$ is the transition relation of $M$ at time steps $i$ and $i+1$. Hence, the equation $\bigwedge^{k-1}_{i=0} R\left(s_{i},s_{i+1}\right)$ represents the set of all executions of $M$ of length $k$. $\neg \phi_{k}$ represents the condition that $\phi$ is violated in state $k$, which is reached by a bounded execution of $M$ of length $k$. Finally, the resulting (bit-vector) equation is translated to conjunctive normal form in linear time and passed to a SAT solver for checking satisfiability. Eq.~(\ref{bounded-model-checking-biere}) can be used to check safety properties~\cite{PrasadBG05}. Liveness properties ({\it e.g.}, starvation, deadlock) that contain the Linear-time Temporal Logic (LTL) operator $F$ are checked by encoding $\neg \phi_{k}$ in a loop within a bounded execution of length at most $k$, such that $\phi$ is violated on each state in the loop. In this case, Eq.~\ref{bounded-model-checking-biere} can be rewritten as:

@@

The ternary operator $f \: \: ? \: \: t_1 : t_2$, shown in Figure~\ref{figure:verification-condition}(b), is converted into the conditional expression $\mathit{ite}(f, t_1, t_2)$ that takes as its first argument the Boolean formula $f$ and, depending on its value, selects either the second (i.e., $t_1$) or the third argument (i.e., $t_2$). In order to verify that the assertion $a<=N$ holds, its negation is added to $\psi$ and it is checked whether the entire formula is satisfiable, through an off-the-self SMT solver. Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can be further transformed into a (equisatisfiable) conjunctive-normal-form formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\mathcal{NP}}$-complete problem~\cite{PatarinG97}).

@@

For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}. If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations. Nonetheless, this invariant is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we initially have $x=0.9$, then $x<0$ after the fourth iteration. As a consequence, even if invariant generation procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that \textit{k}-induction verifiers can automatically prove program correctness. In this specific example, an inductive invariant would be $x>1$, given that if we can check that if $x>1$ before the initial iteration, then also $x>1$ after \textit{k} iterations.
%
\begin{figure}[h]
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
#include <assert.h>
int main(void) {
 float x=2;
 while (1) {
   x = ((2*x) - 1);
 }
 return 0;
}
\end{lstlisting}
\end{minipage}
\caption{Motivating example for inductive invariants.}
\label{figure:inductive-invariant}
\end{figure}

@@

Nonetheless, there is~little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in ECPS verification. That happens due to the distinctive characteristics mentioned earlier, which influence ECPS development and also verification processes. Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination and configuration of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants related to memory and variable aliasing~\cite{Bradley07}.

>>>>

Bounded Model Checking (BMC) based on SAT was originally proposed to verify hardware designs~\cite{Biere99,handbook09}. Indeed, Biere {\it et al.} were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by standard SAT solvers. BMC based on SMT~\cite{BarrettSST09}, in turn, was originally proposed to deal with increasing software verification complexity~\cite{Armando2009}. 
%
In general, BMC techniques aim to check the violation of a given (safety) property at a given system depth, as shown in Figure~\ref{bounded-model-checking}. 
%
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{figure2.jpg}
	\caption{Bounded Model Checking.}
	\label{bounded-model-checking}
\end{figure}

@@

\noindent Here, $\phi_{k}$ represents a safety property $\phi$ in step $k$, $I$ is the set of initial states of $M$, $R\left(s_{i},s_{i+1}\right)$ is the transition relation of $M$ at time steps $i$ and $i+1$. Hence, the equation $\bigwedge^{k-1}_{i=0} R\left(s_{i},s_{i+1}\right)$ represents the set of all executions of $M$ of length $k$. $\neg \phi_{k}$ represents the condition that $\phi$ is violated in state $k$, which is reached by a bounded execution of $M$ of length $k$. Finally, the resulting (bit-vector) equation is translated to conjunctive normal form in linear time and passed to a SAT solver for checking satisfiability. Eq.~(\ref{bounded-model-checking-biere}) can be used to check safety properties~\cite{PrasadBG05}. Liveness properties ({\it e.g.}, starvation, deadlock) that contain the Linear-time Temporal Logic (LTL) operator $F$ are checked by encoding $\neg \phi_{k}$ in a loop within a bounded execution of length at most $k$, such that $\phi$ is violated on each state in the loop~\cite{JMorse15}. In this case, Eq.~\ref{bounded-model-checking-biere} can be rewritten as:

@@

The ternary operator $f \: \: ? \: \: t_1 : t_2$, shown in Figure~\ref{figure:verification-condition}(b), is converted into the conditional expression $\mathit{ite}(f, t_1, t_2)$ that takes as its first argument the Boolean formula $f$ and, depending on its value, selects either the second (i.e., $t_1$) or the third argument (i.e., $t_2$). In order to verify that the assertion $a<=N$ holds, its negation is added to $\psi$ and it is checked whether the entire formula is satisfiable, through an off-the-self SMT solver. Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can be further transformed into a (equisatisfiable) conjunctive-normal-form formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\mathcal{NP}}$-complete problem~\cite{PatarinG97}.

@@

For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}. If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations. Nonetheless, this invariant is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we initially have $x=0.9$, then $x<0$ after the fourth iteration. As a consequence, even if invariant generation procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that \textit{k}-induction verifiers can automatically prove program correctness. In this specific example, an inductive invariant would be $x>1$, given that if $x>1$ before the initial iteration, then also $x>1$ after \textit{k} iterations.
%
\begin{figure}[h]
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
#include <assert.h>
int main(void) {
 float x=2;
 while (1) {
   x = ((2*x) - 1);
 }
 return 0;
}
\end{lstlisting}
\end{minipage}
\caption{Motivating example for inductive invariants.}
\label{figure:inductive-invariant}
\end{figure}

@@

Nonetheless, there is~little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in ECPS verification and synhesis. That happens due to the distinctive characteristics mentioned earlier, which influence ECPS development and also verification processes. Additionally, there is still a lack of studies for software verifiers to exploit the combination and configuration of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants related to memory and variable aliasing~\cite{Bradley07}.
